\begin{description}
  \item[Namespaces:] InputSystem
  \item[Headers:] IInputListener.h, InputActions.h, InputMgr.h, KeyCodes.h, OISListener.h
  \item[Source files:] InputActions.cpp, InputMgr.cpp, OISListener.cpp
  \item[Classes:] InputMgr, IInputListener, OISListener
  \item[Libraries used:] OIS
\end{description}

\section{Purpose of the input system}

Since each game must somehow react to the input from the player, the input system was implemented which can handle keyboard and mouse events and forward them to other systems.

In the following section the way of receiving and distributing input events is described.

\section{Input manager}
The needs of games are different, but the ways they want to access the input devices are still the same. Either they want to receive a notification when something interesting happens or they want to poll the device for its current state. The \emph{InputSystem::InputMgr} class implements both of the approaches.

The first one is provided via the \emph{InputSystem::IInputListener} interface which should be implemented and then registered by using the \emph{InputMgr::\-AddInputListener} method for receiving the notification in callbacks of the interface.

The second approach is supported by multiple methods for querying the device state. The \emph{InputMgr::IsKeyDown} method returns whether any specific key is currently held down while the \emph{InputMgr::IsMouseButtonPressed} method does the same with the currently pressed mouse button. Finally the \emph{InputMgr::GetMouseState} returns a whole bunch of mouse related information such as a cursor and wheel position or pressed buttons.

To keep things synchronized the event processing is executed in the main game thread. At the beginning of each iteration of the game loop the \emph{InputMgr::Capture} method is called where the events are recognized and then distributed to the listeners.

The \emph{InputSystem::InputMgr} class is only a proxy class which calls methods of the implementation specific class \emph{InputSystem::OISListener} which implements the \emph{OIS::MouseListener} and \emph{OIS::KeyListener} interfaces from the OIS library used for the platform independent input management.


%\section{Glossary}
%This is a glossary of the most used terms in the previous sections:
%
%\begin{description}
%  \item[Resource] -- a unit of data the game will be working with as a whole. The data is usually stored in an external device.
%\end{description}


%\begin{thebibliography}{9}                                                                                                
%\bibitem {angelscript}AngelScript documentation, file /AngelScript/index.html
%\end{thebibliography}
